---
title: "Final Code- Bloody IMPACT Analysis"
author: "Paola Fonseca"
date: "2023-10-24"
output: html_document
---
rm(list=ls()) #this clears the workspace
graphics.off(); #close all graphics windows


##Done Installing the IMPACT data set based on Pavia Study:
```{r Insalling and loading data set}
#####Load the data from the Copy_of_20161005_PHC_Data_Merge

# install.packages("readxl")
library(readxl)

##Use when used laptop
#raw_impact <- read_excel("/Users/paolafonsecaromero/Desktop/Blood_Pathogen_project/Pavia Study-impact/Copy of 20161005 PCH Data Merge.xlsx")
#Gave me a warning of discrepancies suggesting data cleaning and handling before analysis. 

#Used when using Desktop
impact_raw <- read_excel("~/Desktop/Blood_Pathogen_project/Pavia Study-impact/Copy of 20161005 PCH Data Merge.xlsx", 
  sheet = "Data")

#impact_raw <- read_excel("C:/Users/u6030586/Box/Documents/all transfers complete/IMPACT - done done/Copy of 20161005 PCH Data Merge.xlsx", 
#    sheet = "Data")

View(impact_raw)
```

##done Creating fix data set and adding patient exclusion criterias
```{r Created a fix data set and handeling; adding exclusion criterias (which pavia preformed in his paper)in this data set}
library(dplyr)

# Create the "fix" data frame to avoid in modifying the original data.
fix <- impact_raw

#Based on Figure.1 on the Pavia study they have exuded the following: Lost to follow up, no specimen returned and excluded before enrollment.
#After observation on data dictionary the variable "Subjectstatus" have these response and also have the same number has the Pavia study.
# table(impact_raw$SubjectStatus) #1157 patients were enrolled, 8patients were excluded before enrollment, 137 lost to follow up and 234 patients had no specimen returned.

#Selecting the patients that have as a value "Enrolled" in the "Subjectstatus" column.
fix <- fix %>%
  filter(SubjectStatus == "Enrolled") #Leaving me with 1,157 patients.

```

##done Recode variables such as: Age group, race, ethnicity and season
```{r Recoded variables}

fix <- fix %>%
  mutate(subjectagecat = ifelse(SUBJECTAGE < 0.5, 1, ifelse(SUBJECTAGE < 2, 2, ifelse(SUBJECTAGE < 5, 3, ifelse(SUBJECTAGE < 12, 4, 5)))) ) %>%
  mutate(subjectagecat2 = ifelse(subjectagecat %in% c(1, 2), 1, ifelse(subjectagecat %in% c(3), 2, ifelse(subjectagecat %in% c(4, 5), 3, NA)))
  )

#subjectagecat:
#less than 0.5, it assigns the value 1.
# age between 0.5 and less than 2, it assigns the value 2.
# age between 2 and less then 5, it assigns the value 3.
# age between 5 and less than 12, it assigns the value 4.
# age is 12 or greater, it assigns the value 5.

#Subjectagecat2:
# if subject is either 1 or 2 the corresponding age categories less then 2, it assigns the value to 1
# if subjectage is 3 corresponding age category 2-5 it assigns the value 2
#if subject is either 4 or 5 corresponding to age categorie 5-12 and 12+ it asaigns the value 3
#if subject does not match any of these conditions, it asigns NA.

# table(fix$subjectagecat)
# table(fix$subjectagecat2)
# #Mean/media calculations
# summary(fix$subjectagecat)
# summary(fix$subjectagecat2)

#Race distribution:
fix$RACE <- ifelse(fix$RACE == "White", 1,
                  ifelse(fix$RACE == "Black or African American", 2, 3))
# table(fix$RACE)

#Ethnicity distribution:
fix$ETHNICITY <- ifelse(
 fix$ETHNICITY == "Hispanic or Latino", 1,
  ifelse(fix$ETHNICITY == "Non-hispanic or Latino", 2,
    3  # Default value for other cases
  )
)
# table(fix$ETHNICITY)


#Season
#Extracting the month and year from the ONSET columns
library(lubridate)
fix$onset_month <- as.integer(format(fix$ONSET, "%m"))
fix$onset_year <- as.integer(format(fix$ONSET, "%Y"))

#Used if else statements to categorize the season based on onset_month
fix$season <- ifelse(fix$onset_month %in% c(1, 2, 3), 1, #winter 261
                     ifelse(fix$onset_month %in% c(4, 5, 6), 2, #spring 233
                            ifelse(fix$onset_month %in% c(7, 8, 9), 3, #summer 371
                                   ifelse(fix$onset_month %in% c(10, 11, 12), 4, NA)))) #fall 292
# table(fix$season)
# table(fix$onset_month) #Confirm, notice that have different distribution in season's compared with pavia's.

###I used this with the fix data frame that contains all the patients after exclusion of the variables in Pavia.table1.



```

#done EPEC detection exclusion, fixing STEC O157 non O157
```{r Observation of distribution and handeling of EPEC}

#For C.Diff I will remove the patients younger than 2 years old + positive and remove all patients with EPEC.

# #Observation of pathogen column EPEC and C.diff before changes
# fix_na_etiology_counts <- colSums(is.na(fix)) # Pathogen columns have 196 NA's 
# table(fix$`EPECInterpretation (D/ND)`) #In the fix data set we have 143 D, 28 NA and 790 ND
# table(fix$`C. difficile Interpretation (D/ND)`) #In the fix data set we have 137 D and 824 ND

#Remove from the fix data set the column that have EPEC.
#I belive this approach is more effective since it removes detection of the bacteria and not the the patients, since some patients that have EPEC could have another pathogen.
fix_no_EPEC <- fix %>%
  select(-`EPECInterpretation (D/ND)`)
# str(fix_no_EPEC) #EPEC column removed

#Observing the distribution of the age detection within the C.diffile column.

# table(fix_no_EPEC$`C. difficile Interpretation (D/ND)` , fix_no_EPEC$subjectagecat)
#          <6 months  6-23 months       2-4 years   5-11 years    12-17 years
#  D          19           70                 18          18           12
#  ND        78          235                 207          208          96

#There are 89 patients in the fix data set that have detection of C.diff that are younger than 2 years. I will not edit this column, I will create a new column that contains "true positives" of the patients that are  =<2 years old in this column the 89 patients will have 0. 
fix_no_EPEC_C.diffnew <- fix_no_EPEC  %>%
  mutate(C.diff.new = ifelse(`C. difficile Interpretation (D/ND)` == "D" & subjectagecat %in% c(3, 4, 5), 1,
                            ifelse(`C. difficile Interpretation (D/ND)` == "D" & subjectagecat %in% c(1, 2), 0,
                                   ifelse(`C. difficile Interpretation (D/ND)` == "ND", 0, NA))))
#Confirming that patients that are younger than 2 years have 0 in c.diff new column
# table(fix_no_EPEC_C.diffnew$C.diff.new, fix_no_EPEC$subjectagecat) # Shows that patients that have detection of c.diff in age group <6 months and 6-23 months are 0.
  #     1   2   3   4   5
  # 0  97 305 207 208  96
  # 1   0   0  18  18  12


###Non-0157 variable Implementation
# table(fix_no_EPEC_C.diffnew$`E. coli O157 Interpretation (D/ND)`) #E.coli has 6 D, 933 NA and 22 ND
# table(fix_no_EPEC_C.diffnew$`STECInterpretation (D/ND)`) #STEC has 28 D and 933 ND


#Creating variable variable for Non-o157, based on the E.coli and STEC information.
#per Daniel for sample size: STEC - 0157 = non0157
fix_no_EPEC_C.diffnew_non0157 <- fix_no_EPEC_C.diffnew %>%
  mutate(Non_O157 = ifelse(`E. coli O157 Interpretation (D/ND)` == "D" & `STECInterpretation (D/ND)` == "D", "ND",
                     ifelse(`E. coli O157 Interpretation (D/ND)` == "ND" & `STECInterpretation (D/ND)` == "D", "D", "ND")))

#Confirming that Non_0157 gave me the correct calculation
table(fix_no_EPEC_C.diffnew_non0157$Non_O157)
table(fix_no_EPEC_C.diffnew_non0157$C.diff.new)

# temp <- as.data.frame(cbind(fix_no_EPEC_C.diffnew_non0157$`E. coli O157 Interpretation (D/ND)`,
#                             fix_no_EPEC_C.diffnew_non0157$Non_O157,
#                             fix_no_EPEC_C.diffnew_non0157$`STECInterpretation (D/ND)`))

#NOTE TO SELF: fix_no_EPEC_C.diffnew_non0157 IS THE NEW DATASET THAT HAS CDIFF AND STECO157 FIXED


```
###Converting the fix_no_EPEC_C.diffnew_non0157 pathogen results to numeric 
```{r}
#In the pathogen columns (seen below), there are 196 NA counts, there are also variable types such as D for detection, ND for not detected and string NA (NA character in the excel file) for E.coli. 

new_pathogen_columns <- c("Adenovirus F Interpretation (D/ND)", "Campylobacter Interpretation (D/ND)","C.diff.new","Cryptosporidium Interpretation (D/ND)", "C. cayetanensis Interpretation (D/ND)", "ETEC Interpretation (D/ND)", "E. coli O157 Interpretation (D/ND)", "EAECInterpretation (D/ND)", "STECInterpretation (D/ND)", "E. histolytica Interpretation (D/ND)", "G. lamblia Interpretation (D/ND)", "Astrovirus Interpretation (D/ND)", "Norovirus Interpretation (D/ND)", "P. shigelloides Interpretation (D/ND)", "Rotavirus A Interpretation (D/ND)", "Salmonella Interpretation (D/ND)", "Sapovirus Interpretation (D/ND)", "Shigella/EIEC Interpretation (D/ND)", "Vibrio Interpretation (D/ND)", "V. cholerae Interpretation (D/ND)", "Y. enterocolitica Interpretation (D/ND)", "Non_O157")

#Verify na counts again
# na_counts_per_column <- colSums(is.na(fix_no_EPEC_C.diffnew_non0157[, new_pathogen_columns]))
# print(na_counts_per_column)

#Because I have ND and NA string values in E.coli only, I will convert the NA and ND to 0, without modifying the 196 NA values.

#Patients PCR responses had ND and D and E.coli had sting NA also. 
  #only typed in NA is converted to 0, truly empty cells remain empty
fix_clean <- fix_no_EPEC_C.diffnew_non0157 %>%
  mutate_at(vars(all_of(new_pathogen_columns)), ~ifelse(. == "D", 1, ifelse(. %in% c("ND", "NA"), 0, .)))
#note to self: this "NA" is a typed in NA, not a truly empty cell

# View(fix_clean)
# str(fix_clean)
# table(is.na(fix_no_EPEC_C.diffnew_non0157$`V. cholerae Interpretation (D/ND)`))
# table(is.na(fix_clean$`V. cholerae Interpretation (D/ND)`))

#Convert the selected columns to numeric
fix_clean_final <- fix_clean %>%
 mutate(across(all_of(new_pathogen_columns), ~ as.numeric(.)))

# View(fix_clean_final)
# str(fix_clean_final)

#Display the distribution of values in each pathogen column after converting them to numeric.
# for (col in new_pathogen_columns) {
#   cat("Column:", col, "\n")
#   print(table(fix_clean_final[[col]]))
#   cat("\n")
# }


# create a single column that lists all the pathogens detected
fix_clean_final$pathogens_detected <- NA
for (i in 1:nrow(fix_clean_final)) {
  #get name of pathogens detected for each row
  detected_pathogens <- colnames(fix_clean_final[new_pathogen_columns])[fix_clean_final[new_pathogen_columns][i, ] == 1]
  fix_clean_final$pathogens_detected[i] <- paste(detected_pathogens, collapse=", ")

}

#Create a new columns that contains the count of detected pathogen in each row
fix_clean_final$detected_pathogen_count <- rowSums(fix_clean_final[,new_pathogen_columns[!new_pathogen_columns %in% c("Non_O157","E. coli O157 Interpretation (D/ND)")]])
# #Created summary table of the counts of detected pathogens.
#excluding non_o157 and o157 from the row sum, only including STEC in the row sum since otherwise redundant
# bloody_pathogens_final <- table(fix_blo_mPCR$detected_pathogen_count)
# print(bloody_pathogens_final) #42 patients had 0 in their rows, 43 had 1 detection, 20 had two detentions, 5 detection three detentions and 1 had five detection.


```

###Subset used in the analysis:
```{r}
#Subseting patients with DIARRHEA, including w/o PCR and w/o data in blood in stool ~ 1020 patients/rows)}
fix_Diarrhea<- fix_clean_final %>%
  filter(DIARRHEA == "Yes") 

#Subseting patients with Bloody diarrhea (with and without PCR) n = 126
fix_bloody<- fix_clean_final %>%
  filter(DIA_BLOODY == "Yes") 

#Subsetting patients w/o bloody diarrhea (with and without PCR) n = 860
fix_non_bloody <- fix_Diarrhea %>% 
  filter(DIA_BLOODY != "Yes")

#Subsetting patients with bloody diarrhea with PCR only single pathogen present
blo_one_pathogen_detected <- fix_Diarrhea %>%
  filter(DIA_BLOODY == "Yes" & detected_pathogen_count == 1)

#Subsetting patients with Bloody diarrhea with PCR information n= 111
fix_blo_mPCR <- fix_Diarrhea %>% 
  filter(DIA_BLOODY == "Yes" & !is.na(`Shigella/EIEC Interpretation (D/ND)`))

#dataset for additional wrangling for abx supplemental table
fix_blo_abx <- fix_blo_mPCR

#Subsetting patient w/o bloody diarrhea with PCR information N = 731
fix_non_blo_mPCR <- fix_Diarrhea %>% 
  filter(DIA_BLOODY != "Yes" & !is.na(`Shigella/EIEC Interpretation (D/ND)`))

#Subsetting patients with and without bloody diarrhea who have PCR N=842
bloody_and_non_bloody <- fix_Diarrhea %>%
  filter(!is.na(DIA_BLOODY) & !is.na(`Shigella/EIEC Interpretation (D/ND)`))

#subsetting to all patients with PCR, with bloody and non-bloody N=842
fix_mPCR <- fix_clean_final %>%
  filter(DIARRHEA == "Yes" & !is.na(DIA_BLOODY) & !is.na(`Shigella/EIEC Interpretation (D/ND)`))


```



```{r}
table(is.na(fix_Diarrhea$SOCIAL_ANIMAL))
```


###Supplement Table.1 - All Diarrhe patients (including without PCR and wihout data on blood in stool)
```{r Convertion of D/ND to numeric}

#Selecting specific columns and created new data set
fix_diarrhea_char<- fix_Diarrhea %>%
  select(SEX, RACE, ETHNICITY, subjectagecat, COND_NONE, COND_IMMUNO, INSURANCE, SOCIAL_SCHOOL, SOCIAL_ANIMAL, SOCIAL_INTLTRAVEL, season)

library(gtsummary)

fix_diarrhea_char %>%
  tbl_summary(
    missing = "no"
  )


```

#Bloody All (with and without PCR)
```{r Subseting patients with Bloody diarrhea (including without PCR) N=126 }

#Selecting specific columns and created new data set
fix_bloody_char<- fix_bloody %>%
  select(SEX, RACE, ETHNICITY, subjectagecat, COND_NONE, COND_IMMUNO, INSURANCE, SOCIAL_SCHOOL, SOCIAL_ANIMAL, SOCIAL_INTLTRAVEL, season)

tbl_summary(fix_bloody_char)

```


#Non-Bloody All (with and without PCR)
```{r}
###From the 1020, I have patients that have blood diarrhea. Due to that I have to remove them.

#Selecting specific columns and created new data set
fix_nonblo_character <- fix_non_bloody%>%
  select(SEX, RACE, ETHNICITY, subjectagecat, COND_NONE, COND_IMMUNO, INSURANCE, SOCIAL_SCHOOL, SOCIAL_ANIMAL, SOCIAL_INTLTRAVEL, season)

tbl_summary(fix_nonblo_character)

```

#Bloody with PCR information (n=111)
```{r Subseting patients with bloody diarrhea with PCR}
#Created a data set with only the new pathogen,subject categories and subject ID(to merge the other patients characteristics).

fix_blo_mPCR_char <- fix_blo_mPCR %>%
  select(SEX, RACE, ETHNICITY, subjectagecat, COND_NONE, COND_IMMUNO, INSURANCE, SOCIAL_SCHOOL,
         SOCIAL_ANIMAL,SOCIAL_INTLTRAVEL, season) 

#Table
tbl_summary(fix_blo_mPCR_char)

```

#Non-Bloody with PCR information (n=731)
```{r}

fix_non_blo_mPCR_char <- fix_non_blo_mPCR %>% select(SEX, RACE, ETHNICITY, subjectagecat, COND_NONE, COND_IMMUNO, INSURANCE,
                                                     SOCIAL_SCHOOL, SOCIAL_ANIMAL,SOCIAL_INTLTRAVEL, season, SUBJECTAGE)

#Table
tbl_summary(fix_non_blo_mPCR_char)

```

#Mean/Media calculations of these subsets
```{r}
#Diarrhea all n=1020
table(is.na(fix_Diarrhea$SUBJECTAGE)) #No missing values
mean(fix_Diarrhea$SUBJECTAGE) #4.7
median(fix_Diarrhea$SUBJECTAGE) #2.5

# Bloody Diarrhea all n=126
table(is.na(fix_bloody$SUBJECTAGE)) #No missing values
mean(fix_bloody$SUBJECTAGE) #6.5
median(fix_bloody$SUBJECTAGE) #5.31

#Non-Bloody All n=860
table(is.na(fix_non_bloody$SUBJECTAGE)) #No missing values
mean(fix_non_bloody$SUBJECTAGE) #4.4
median(fix_non_bloody$SUBJECTAGE) #2.2

#Bloody with PCR  n=111
table(is.na(fix_blo_mPCR$SUBJECTAGE)) #No missing values
mean(fix_blo_mPCR$SUBJECTAGE) #6.5
median(fix_blo_mPCR$SUBJECTAGE) #5.2

#Non-Bloody with PCR n=731
table(is.na(fix_non_blo_mPCR$SUBJECTAGE)) #No missing values
mean(fix_non_blo_mPCR$SUBJECTAGE) #4.3
median(fix_non_blo_mPCR$SUBJECTAGE) #2.25


```


#Table 1: Frequency of detection of the 21 pathogens in the PCR, among patients presenting with bloody diarrhea, stratified by age group.Co-occuring pathogens are listed multiple times.
```{r}

fix_blo_mPCR_fortable <- fix_blo_mPCR %>% select(`Adenovirus F Interpretation (D/ND)`, `Campylobacter Interpretation (D/ND)`, 
                                                 C.diff.new, `STECInterpretation (D/ND)`,
                                                 `Cryptosporidium Interpretation (D/ND)`, `C. cayetanensis Interpretation (D/ND)`, 
                                                 `ETEC Interpretation (D/ND)`, `E. coli O157 Interpretation (D/ND)`, `EAECInterpretation (D/ND)`, 
                                                 `E. histolytica Interpretation (D/ND)`, `G. lamblia Interpretation (D/ND)`, 
                                                 `Astrovirus Interpretation (D/ND)`, `Norovirus Interpretation (D/ND)`, 
                                                 `P. shigelloides Interpretation (D/ND)`, `Rotavirus A Interpretation (D/ND)`, 
                                                 `Salmonella Interpretation (D/ND)`, `Sapovirus Interpretation (D/ND)`, 
                                                 `Shigella/EIEC Interpretation (D/ND)`, `Vibrio Interpretation (D/ND)`, 
                                                 `V. cholerae Interpretation (D/ND)`, `Y. enterocolitica Interpretation (D/ND)`, 
                                                 Non_O157, subjectagecat)

#Used the tbl.summary to create table
tbl_summary(fix_blo_mPCR_fortable)

fix_blo_mPCR_fortable %>%
  tbl_summary(by = subjectagecat)
```

```{r}
#Count of each pathogen in pathogen detection among the bloody diarrhea patients.

#Bacteria

library(dplyr)

#Bacteria
fix_bacteria_mPCR <- fix_blo_mPCR_fortable %>%
                                           select(`Campylobacter Interpretation (D/ND)`, 
                                                 C.diff.new, `STECInterpretation (D/ND)`,
                                                 `Cryptosporidium Interpretation (D/ND)`, 
                                                 `ETEC Interpretation (D/ND)`, `E. coli O157 Interpretation (D/ND)`, `EAECInterpretation (D/ND)`, 
                                                 `P. shigelloides Interpretation (D/ND)`,
                                                 `Salmonella Interpretation (D/ND)`, 
                                                 `Shigella/EIEC Interpretation (D/ND)`, `Vibrio Interpretation (D/ND)`, 
                                                 `V. cholerae Interpretation (D/ND)`, `Y. enterocolitica Interpretation (D/ND)`, 
                                                 Non_O157) %>%
  rowwise() %>%
  mutate(detections = sum(c_across(everything()), na.rm = TRUE)) %>%
  ungroup()
  
table(fix_bacteria_mPCR$detections)
tbl_summary(fix_bacteria_mPCR) #We have 53 detentions on children with bloody diarrhea which has co-occurring pathogens.

#Virus
fix_virus_mPCR <- fix_blo_mPCR_fortable %>%
                                           select(`Adenovirus F Interpretation (D/ND)`,
                                                 `Astrovirus Interpretation (D/ND)`, `Norovirus Interpretation (D/ND)`, `Rotavirus A Interpretation (D/ND)`, `Sapovirus Interpretation (D/ND)`) %>%
  rowwise() %>%
  mutate(detections = sum(c_across(everything()), na.rm = TRUE)) %>%
  ungroup()
  
table(fix_virus_mPCR$detections)
tbl_summary(fix_virus_mPCR)

#Protozoa
fix_protozoa_mPCR <- fix_blo_mPCR_fortable %>%
                                           select(`Cryptosporidium Interpretation (D/ND)`, `C. cayetanensis Interpretation (D/ND)`,
                                                 `E. histolytica Interpretation (D/ND)`, `G. lamblia Interpretation (D/ND)`) %>%
  rowwise() %>%
  mutate(detections = sum(c_across(everything()), na.rm = TRUE)) %>%
  ungroup()
  
table(fix_protozoa_mPCR$detections)
tbl_summary(fix_protozoa_mPCR)

#No detentions n=42 
fix_no_detections_mPCR <- fix_blo_mPCR_fortable %>%
  select(-subjectagecat) %>%
 rowwise() %>%
  mutate(detections = sum(c_across(everything()), na.rm = TRUE)) %>%
  ungroup()

table(fix_no_detections_mPCR$detections == 0)

```


#Supplement table 2. Frequency table for each of the 21 pathogens tested in the PCR panel, excluding patients with multiple pathogens.
```{r}
blo_one_pathogen_detected_fortable <- blo_one_pathogen_detected %>% 
  select(`Adenovirus F Interpretation (D/ND)`, `Campylobacter Interpretation (D/ND)`, 
                                                 C.diff.new, `STECInterpretation (D/ND)`,
                                                 `Cryptosporidium Interpretation (D/ND)`, `C. cayetanensis Interpretation (D/ND)`, 
                                                 `ETEC Interpretation (D/ND)`, `E. coli O157 Interpretation (D/ND)`, `EAECInterpretation (D/ND)`, 
                                                 `E. histolytica Interpretation (D/ND)`, `G. lamblia Interpretation (D/ND)`, 
                                                 `Astrovirus Interpretation (D/ND)`, `Norovirus Interpretation (D/ND)`, 
                                                 `P. shigelloides Interpretation (D/ND)`, `Rotavirus A Interpretation (D/ND)`, 
                                                 `Salmonella Interpretation (D/ND)`, `Sapovirus Interpretation (D/ND)`, 
                                                 `Shigella/EIEC Interpretation (D/ND)`, `Vibrio Interpretation (D/ND)`, 
                                                 `V. cholerae Interpretation (D/ND)`, `Y. enterocolitica Interpretation (D/ND)`, 
                                                 Non_O157, subjectagecat)


##TABLE HERE
blo_one_pathogen_detected_fortable %>%
  tbl_summary(by = subjectagecat)
```

```{r}
#Total count of each pathogen in single detection among the bloody diarrhea patients.

#Bacteria n=45, 41% bacteria
table(blo_one_pathogen_detected_fortable$`Campylobacter Interpretation (D/ND)`) 
# 0  1 
#48  4 
table(blo_one_pathogen_detected_fortable$C.diff.new)
# 0  1 
#46  6 
table(blo_one_pathogen_detected_fortable$`ETEC Interpretation (D/ND)`)
# 0 
#52
table(blo_one_pathogen_detected_fortable$`EAECInterpretation (D/ND)`)
# 0  1 
#50  2 
table(blo_one_pathogen_detected_fortable$`Salmonella Interpretation (D/ND)`)
# 0  1 
#48  4 
table(blo_one_pathogen_detected_fortable$`P. shigelloides Interpretation (D/ND)`)
#0 
#52 
table(blo_one_pathogen_detected_fortable$`Vibrio Interpretation (D/ND)`)
#0 
#52
table(blo_one_pathogen_detected_fortable$`STECInterpretation (D/ND)`)
#0  1 
#43  9
table(blo_one_pathogen_detected_fortable$Non_O157)
#0  1 
#47  5 
table(blo_one_pathogen_detected_fortable$`E. coli O157 Interpretation (D/ND)`)
#0  1 
#48  4 
table(blo_one_pathogen_detected_fortable$`Shigella/EIEC Interpretation (D/ND)`)
# 0  1 
#41 11 
table(blo_one_pathogen_detected_fortable$`Y. enterocolitica Interpretation (D/ND)`)
#0 
#52
table(blo_one_pathogen_detected_fortable$`V. cholerae Interpretation (D/ND)`)
#0 
#52

#Viruses n=14, 13% viruses
table(blo_one_pathogen_detected_fortable$`Adenovirus F Interpretation (D/ND)`)
# 0 
#52
table(blo_one_pathogen_detected_fortable$`Astrovirus Interpretation (D/ND)`)
# 0  1 
#50  2
table(blo_one_pathogen_detected_fortable$`Norovirus Interpretation (D/ND)`)
# 0  1 
#48  4 
table(blo_one_pathogen_detected_fortable$`Sapovirus Interpretation (D/ND)`)
#0  1 
#48  4 
table(blo_one_pathogen_detected_fortable$`Rotavirus A Interpretation (D/ND)`)
# 0  1 
#48  4

#Protozoa n=2, 2% protozoa
table(blo_one_pathogen_detected_fortable$`G. lamblia Interpretation (D/ND)`)
# 0  1 
#50  2 
table(blo_one_pathogen_detected_fortable$`Cryptosporidium Interpretation (D/ND)`)
# 0 
#52
table(blo_one_pathogen_detected_fortable$`C. cayetanensis Interpretation (D/ND)`)
#0 
#52
table(blo_one_pathogen_detected_fortable$`E. histolytica Interpretation (D/ND)`)
#0 
#52
```


#Data Wrangeling for abx analysis:
```{r}
#Paola & Sharia start over on this section together
#these variables are asking about meds before presenting to ED
# ANTIBIOTICS = "Did your child take any antibiotic medications (e.g., ciprofloxacin, levofloxacin, trimethoprim-sulfamethoxazole, flagyl) for this illness before coming to the hospital?"
# ANTIBIOTICS_OTHER = "Did your child take any other medications for this illness before coming to the hospital?"
#_LIST for both of these are listed the actual meds taken before presenting to care

### Create new columns to indicate antibiotics administered during and after ED
fix_blo_abx$Abx_start_in_ED <- NA
fix_blo_abx$Abx_start_after_ED <- NA

# Loop through each row of the dataset
for (i in 1:nrow(fix_blo_abx)) {
    # Create vectors to store antibiotic names for during and after ED administrations
    abx_start_in_ed <- character()
    abx_start_after_ed <- character()
    
    # Loop through each MED variable - these are the MED variables which are abx based on data dictionary
    for (j in c(5, 6, 7, 8, 9, 10, 11, 16)) {
        med_start <- fix_blo_abx[[paste0("MED", j, "_START")]][i]
        med_stop <- fix_blo_abx[[paste0("MED", j, "_STOP")]][i]
        med_name <- fix_blo_abx[[paste0("MED", j)]][i]
        ed_discharge_date <- fix_blo_abx$DISCHARGE_ED_DATE[i]
        
        # Check if MED is administered and if abx started during or after ED
        if (!is.na(med_name) && med_name == "Yes" && !is.na(med_start) ) {
            if (med_start <= ed_discharge_date ) {
                abx_start_in_ed <- c(abx_start_in_ed, paste0("MED", j))
            } else if (med_start > ed_discharge_date ) {
                abx_start_after_ed <- c(abx_start_after_ed, paste0("MED", j))
            }
        }
    }
    
    # Assign antibiotic names to the appropriate columns
    if (length(abx_start_in_ed) > 0) {
        fix_blo_abx$Abx_start_in_ED[i] <- paste(abx_start_in_ed, collapse = ", ")
    }
    if (length(abx_start_after_ed) > 0) {
        fix_blo_abx$Abx_start_after_ED[i] <- paste(abx_start_after_ed, collapse = ", ")
    }
}


check_abx <- fix_blo_abx %>% select(SubjectID, MED5, MED5_START, MED5_STOP,
                                    MED6, MED6_START, MED6_STOP,
                                    MED7, MED7_START, MED7_STOP,
                                    MED8, MED8_START, MED8_STOP,
                                    MED9, MED9_START, MED9_STOP,
                                    MED10, MED10_START, MED10_STOP,
                                    MED11, MED11_START, MED11_STOP,
                                    MED16, MED16_START, MED16_STOP,
                                    OTHER_ANTIBIOTIC,
                                    DISCHARGE_ED_DATE, Group,
                                    subjectagecat,
                                    Abx_start_in_ED, Abx_start_after_ED)
```


#Supplement table.3 Pathogens detected in cases of bloody diarrhea who recived antibiotics during or after ED.
```{r}
#Selection of specific columns
abx_analysis_final <- fix_blo_abx %>%
  select(SubjectID, pathogens_detected, Abx_start_in_ED, Abx_start_after_ED, subjectagecat,Group,OTHER_ANTIBIOTIC)

abx_analysis_table <- abx_analysis_final %>%
  filter(!is.na(Abx_start_in_ED) | !is.na(Abx_start_after_ED))
View(abx_analysis_table)
# save(abx_analysis_table, file = "C:/Users/u6030586/Box/Documents/all transfers complete/IMPACT - done done/abx_analysis_table.Rdata")
#load(file = "C:/Users/u6030586/.../abx_analysis_table.Rdata")


```

#ODDS RATIO Table.2 Unadjusted and age group-adjusted Odds Ratio (OR) of association between the presence of pathogen and the occurence of bloody diarrhea

```{r}
#Bloody 
table(fix$`Campylobacter Interpretation (D/ND)`)
#  0   1 
#103   8 
table(fix_blo_mPCR_fortable$C.diff.new)
 #0   1 
#102   9
table((fix_blo_mPCR_fortable$`ETEC Interpretation (D/ND)`))
# 0   1 
#107   4 
table(fix_blo_mPCR_fortable$`EAECInterpretation (D/ND)`)
# 0   1 
#101  10
table(fix_blo_mPCR_fortable$`Salmonella Interpretation (D/ND)`)
#0   1 
#105   6 
table(fix_blo_mPCR_fortable$`P. shigelloides Interpretation (D/ND)`)
 #0   1 
#110   1 
table(fix_blo_mPCR_fortable$`STECInterpretation (D/ND)`)
# 0  1 
#97 14
table(fix_blo_mPCR_fortable$Non_O157)
 # 0   1 
#102   9 
table(fix_blo_mPCR_fortable$`E. coli O157 Interpretation (D/ND)`)
#  0   1 
#106   5
```


```{r}
non_bloody_and_bloody_paathogen <- fix_mPCR %>% 
  select(`Adenovirus F Interpretation (D/ND)`, `Campylobacter Interpretation (D/ND)`, 
                                                 C.diff.new, `STECInterpretation (D/ND)`,
                                                 `Cryptosporidium Interpretation (D/ND)`, `C. cayetanensis Interpretation (D/ND)`, 
                                                 `ETEC Interpretation (D/ND)`, `E. coli O157 Interpretation (D/ND)`, `EAECInterpretation (D/ND)`, 
                                                 `E. histolytica Interpretation (D/ND)`, `G. lamblia Interpretation (D/ND)`, 
                                                 `Astrovirus Interpretation (D/ND)`, `Norovirus Interpretation (D/ND)`, 
                                                 `P. shigelloides Interpretation (D/ND)`, `Rotavirus A Interpretation (D/ND)`, 
                                                 `Salmonella Interpretation (D/ND)`, `Sapovirus Interpretation (D/ND)`, 
                                                 `Shigella/EIEC Interpretation (D/ND)`, `Vibrio Interpretation (D/ND)`, 
                                                 `V. cholerae Interpretation (D/ND)`, `Y. enterocolitica Interpretation (D/ND)`, 
                                                Non_O157, subjectagecat,DIA_BLOODY)

#Table of the descriptive statistics of bloody and non-bloody in TABLE.2
non_bloody_and_bloody_paathogen %>%
  tbl_summary(by = DIA_BLOODY)


```


```{r}

#Convert the outcome DIA_BLOODY to numeric
fix_mPCR <- fix_mPCR %>%
  mutate(DIA_BLOODY_num = ifelse(DIA_BLOODY =="Yes",1,0))
# table(fix_mPCR$DIA_BLOODY)
# #  No Yes 
# # 731 111 
# table(fix_mPCR$DIA_BLOODY2)
# #   0   1 
# # 731 111 
# table(is.na(fix_mPCR$DIA_BLOODY))
# # FALSE 
# #   842 
# table(is.na(fix_mPCR$DIA_BLOODY2))
# # FALSE 
# #   842 


#Fit a logistic regression model using the glm package:

#STEC model
stec_model <- glm(DIA_BLOODY_num ~ `STECInterpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(stec_model$coefficients)[2]
exp(confint((stec_model)))[2,]
#Adj model
stec_adj <- glm(DIA_BLOODY_num ~ `STECInterpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(stec_adj$coefficients)[2]
exp(confint((stec_adj)))[2,]

#E.coli 0157 model
e_coli_o157_model <- glm(DIA_BLOODY_num ~ `E. coli O157 Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(e_coli_o157_model$coefficients)[2]
exp(confint((e_coli_o157_model)))[2,]
#Adj Model
e_coli_o157_adj <- glm(DIA_BLOODY_num ~ `E. coli O157 Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(e_coli_o157_adj$coefficients)[2]
exp(confint((e_coli_o157_adj)))[2,]

#Non_O157 model
non_o157_model <- glm(DIA_BLOODY_num ~ Non_O157, data = fix_mPCR, family = "binomial")
exp(non_o157_model$coefficients)[2]
exp(confint((non_o157_model)))[2,]
#Adj model
non_o157_adj <- glm(DIA_BLOODY_num ~ Non_O157 + subjectagecat, data = fix_mPCR, family = "binomial")
exp(non_o157_adj$coefficients)[2]
exp(confint((non_o157_adj)))[2,]

#Campylobacter model
campylobac_model <- glm(DIA_BLOODY_num ~ `Campylobacter Interpretation (D/ND)`, data=fix_mPCR, family= "binomial")
exp(campylobac_model$coefficients)[2]
exp(confint((campylobac_model)))[2,]
#Adj model
campylobac_adj <- glm(DIA_BLOODY_num ~ `Campylobacter Interpretation (D/ND)` + subjectagecat, data=fix_mPCR, family="binomial")
exp(campylobac_adj$coefficients)[2]
exp(confint((campylobac_adj)))[2,]

#P. shigelloides model
p_shigelloides_model <- glm(DIA_BLOODY_num ~ `P. shigelloides Interpretation (D/ND)`, data=fix_mPCR, family= "binomial")
exp(p_shigelloides_model$coefficients)[2]
exp(confint((p_shigelloides_model)))[2,]
#Adj model
p_shigelloides_adj <- glm(DIA_BLOODY_num ~ `P. shigelloides Interpretation (D/ND)` + subjectagecat, data=fix_mPCR, family= "binomial")
exp(p_shigelloides_adj$coefficients)[2]
exp(confint((p_shigelloides_adj)))[2,]

#C. difficile model
c_diff_model <- glm(DIA_BLOODY_num ~ C.diff.new, data =fix_mPCR, family = "binomial")
exp(campylobac_adj$coefficients)[2]
exp(confint((campylobac_adj)))[2,]
#Aj model
c_diff_adj <- glm(DIA_BLOODY_num ~ C.diff.new + subjectagecat, data =fix_mPCR, family = "binomial")
exp(c_diff_adj$coefficients)[2]
exp(confint((c_diff_adj)))[2,]

#Shigella/EIEC model
shigella_model <- glm(DIA_BLOODY_num ~ `Shigella/EIEC Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(shigella_model$coefficients)[2]
exp(confint((shigella_model)))[2,]
#Adj model
shigella_adj <- glm(DIA_BLOODY_num ~ `Shigella/EIEC Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(shigella_adj$coefficients)[2]
exp(confint((shigella_adj)))[2,]

#EAEC model
eaec_model <- glm(DIA_BLOODY_num ~ `EAECInterpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(eaec_model$coefficients)[2]
exp(confint((eaec_model)))[2,]
#Adj model
eaec_adj <- glm(DIA_BLOODY_num ~ `EAECInterpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(eaec_adj$coefficients)[2]
exp(confint((eaec_adj)))[2,]

#Salmonella model
salmonella_model <- glm(DIA_BLOODY_num ~ `Salmonella Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(salmonella_model$coefficients)[2]
exp(confint((salmonella_model)))[2,]
#Adj model
salmonella_adj <- glm(DIA_BLOODY_num ~ `Salmonella Interpretation (D/ND)` + subjectagecat , data = fix_mPCR, family = "binomial")
exp(salmonella_adj$coefficients)[2]
exp(confint((salmonella_adj)))[2,]


#Rotavirus model
rotavirus_model <- glm(DIA_BLOODY_num ~ `Rotavirus A Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(rotavirus_model$coefficients)[2]
exp(confint((rotavirus_model)))[2,]
#Adj model
rotavirus_adj <- glm(DIA_BLOODY_num ~ `Rotavirus A Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(rotavirus_adj$coefficients)[2]
exp(confint((rotavirus_adj)))[2,]

#G.lamblia model
g_lamblia_model <- glm(DIA_BLOODY_num ~ `G. lamblia Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(g_lamblia_model$coefficients)[2]
exp(confint((g_lamblia_model)))[2,]
#Adj model
g_lamblia_adj <- glm(DIA_BLOODY_num ~ `G. lamblia Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(g_lamblia_adj$coefficients)[2]
exp(confint((g_lamblia_adj)))[2,]

#ETEC
etec_model <- glm(DIA_BLOODY_num ~ `ETEC Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(etec_model$coefficients)[2]
exp(confint((etec_model)))[2,]
#Adj model
etec_adj <- glm(DIA_BLOODY_num ~ `ETEC Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(etec_adj$coefficients)[2]
exp(confint((etec_adj)))[2,]

#Sapovirus
sapovirus_model <- glm(DIA_BLOODY_num ~ `Sapovirus Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(sapovirus_model$coefficients)[2]
exp(confint((sapovirus_model)))[2,]
#Adj model
sapovirus_adj <- glm(DIA_BLOODY_num ~ `Sapovirus Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(sapovirus_adj$coefficients)[2]
exp(confint((sapovirus_adj)))[2,]

#Astrovirus
astrovirus_model <- glm(DIA_BLOODY_num ~ `Astrovirus Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(astrovirus_model$coefficients)[2]
exp(confint((astrovirus_model)))[2,]
#Adj model
astrovirus_adj <- glm(DIA_BLOODY_num ~ `Astrovirus Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(astrovirus_adj$coefficients)[2]
exp(confint((astrovirus_adj)))[2,]

#Norovirus
norovirus_model <- glm(DIA_BLOODY_num ~ `Norovirus Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(norovirus_model$coefficients)[2]
exp(confint((norovirus_model)))[2,]
#Ajd model
norovirus_adj <- glm(DIA_BLOODY_num ~ `Norovirus Interpretation (D/ND)` + subjectagecat, data = fix_mPCR, family = "binomial")
exp(norovirus_adj$coefficients)[2]
exp(confint((norovirus_adj)))[2,]

#Adenovirus.F
adenovirus_model <- glm(DIA_BLOODY_num ~ `Adenovirus F Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")
exp(adenovirus_model$coefficients)[2]
exp(confint((adenovirus_model)))[2,]

#Cryptospridum
cryptospridum_model <- glm(DIA_BLOODY_num ~ `Cryptosporidium Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")

#Y. eneterocolitica
y_enterocolitica_model <- glm(DIA_BLOODY_num ~ `Y. enterocolitica Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")

#vibrio
vibrio_model <- glm(DIA_BLOODY_num ~ `Vibrio Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")

#V. cholera
v_cholera_model <- glm(DIA_BLOODY_num ~ `V. cholerae Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")

#E. histolytica
e_histolytic_model <- glm(DIA_BLOODY_num ~ `E. histolytica Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")

#C. cayetanesis
c_cayatenesis_model <- glm(DIA_BLOODY_num ~ `C. cayetanensis Interpretation (D/ND)`, data = fix_mPCR, family = "binomial")


#This code will fit separate logistic regression models for each etiology and print results including odd ratio and CI. Each model will focus solely on one etiology at a time, which aligns with Sharia's feedback regarding examining single models with single etiologies.
```


